\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2023-10-06 paper.tex

% FIXME abstract needs some rewriting, we're placing too much emphasis
% on the coalescent there.
\begin{abstract}
New developments have made it possible to infer genetic genealogies in
the presence of recombination at scale, enabling many
downstream applications in population and statistical genetics.
The structure representing such recombinant genetic ancestry
is usually referred to as an ancestral recombination graph (ARG),
although there is some confusion about the interpretation and little
agreement on the concrete details.
We propose a concrete definition of ARGs
in terms of genomes and their intervals of genetic inheritance (gARG)
and contrast this with the classical event-based definition (eARG).
The eARG definition is derived from on a stochastic process, 
the coalescent with recombination, 
and we show how the assumptions of that process
place limitations on the patterns of inheritance that can be
represented, which are inadequate for modern large-scale datasets.
The eARG representation also requires complete precision about the ordering
of all recombination events, which we will rarely have sufficient
information to infer.
We show, in contrast, that the gARG definition
can fully capture the richness of modern large-scale datasets,
enables fine-grained
levels of precision about recombination to be represented,
and forms the basis of an efficient computational framework.
\end{abstract}

@@

% What do we do?
In this paper, we attempt to address these problems by
providing a precise definition of an ARG that is general enough to encompass
the output of popular simulation and inference methods.
We propose a definition of genetic genealogy in terms of specific genomes
and their intervals of genetic inheritance, and refer to this as
a ``genome ARG'', or ``gARG''.
This contrasts with the classical Griffiths approach,
based on recording evolutionary events, which we call
an ``event ARG'', or ``eARG''.
We show that the gARG encoding has a range of advantages over the
eARG approach, being flexible, efficient and, in particular,
allowing us to systematically encode uncertainty about the
time ordering of recombination events.
For a recent and more biologically oriented introduction to
the topic, see \citet{lewanski2023era}.

>>>>

% FIXME abstract needs some rewriting, we're placing too much emphasis
% on the coalescent there.
\begin{abstract}
New developments have made it possible to infer genetic genealogies in
the presence of recombination at scale, enabling many
downstream applications in population and statistical genetics.
The structure representing such recombinant genetic ancestry
is usually referred to as an ancestral recombination graph (ARG),
although there is some confusion about the interpretation and little
agreement on the concrete details.
We propose a concrete definition of ARGs
in terms of genomes and their intervals of genetic inheritance (gARG)
and contrast this with the classical event-based definition (eARG).
We show that eARGs are limited in the patterns
of genetic inheritance that can be represented,
and require complete precision about all recombination events.
We show, in contrast, that the gARG definition
can fully capture the richness of modern large-scale datasets,
enables fine-grained
levels of precision about recombination to be represented,
and forms the basis of an efficient computational framework.
\end{abstract}

@@

% What do we do?
In this paper, we attempt to address these problems by
providing a precise definition of an ARG that is general enough to encompass
the output of popular simulation and inference methods.
We propose a definition of genetic genealogy in terms of specific genomes
and their intervals of genetic inheritance, and refer to this as
a ``genome ARG'', or ``gARG''.
This contrasts with the classical Griffiths approach,
based on recording evolutionary events, which we call
an ``event ARG'', or ``eARG''.
We show that the gARG encoding has a range of advantages over the
eARG approach, being flexible, efficient and, in particular,
allowing us to systematically encode uncertainty about the
time ordering of recombination events.
For a recent and more biologically oriented introduction to
ARGs, see \citet{lewanski2023era}.
